\thetitle

### 2013-04-17 tfpie/tfpie.tex
small edits from Katie

As an example, Northeastern University teaches functional programming
in the first semester using \emph{How to Design
  Programs}~\cite{dvanhorn:Felleisen2001How}, followed by
object-oriented programming using \emph{How to Design
  Classes}~\cite{local:htdc} in the second semester.  This sequence
was designed to provide a smooth path for incoming students assumed
only to have a competence in high-school level algebra to reach
proficiency in Java by the end of their first year
\cite{dvanhorn:Felleisen2004Structure}.  It was a major improvement
over the previous Java-first curriculum in terms of student success,
attrition, and preparation for subsequent
courses~\cite{dvanhorn:Proulx2006Design}.  However significant
problems remain; in particular, the second semester course violates
the designers' own principles (as recalled
in~\cite{dvanhorn:Bloch2000Scheme}):
%
\begin{enumerate}
\item \emph{introduce only those language constructs that are necessary to
  teach programming principles}, and
\item \emph{choose a language with as
  few language constructs as possible, and one in which they can be
  introduced one at a time}.
\end{enumerate}

@@

At Northeastern, the College of Computer \& Information Science (CCIS)
requires a four course introductory sequence in the first year.  The
first semester features both a course on discrete mathematics as well
as an introduction to programming following the \emph{How to Design
  Programs} curriculum.  The second semester follows with a course on
object-oriented programming and one featuring formal reasoning about
programs, both on paper and with the ACL2 theorem
prover~\cite{dvanhorn:Kaufmann2000ComputerAided}.

@@

This combination sets the constraints in which we designed our
approach.  Our students begin the course with a firm grasp of
data-driven program design, as well as experience with the tools used
in the first course:

>>>>

As an example, Northeastern University teaches functional programming
in the first semester using \emph{How to Design
  Programs}~\cite{dvanhorn:Felleisen2001How}, followed by
object-oriented programming using \emph{How to Design
  Classes}~\cite{local:htdc} in the second semester.  This sequence
was designed to provide a smooth path for incoming students assumed
only to have a competence in high-school level algebra to reach
proficiency in Java by the end of their first year
\cite{dvanhorn:Felleisen2004Structure}.  It was a major improvement
over the previous Java-first curriculum in terms of student success,
attrition, and preparation for subsequent
courses~\cite{dvanhorn:Proulx2006Design}.  However, significant
problems remain; in particular, the second semester course violates
the designers' own principles (as recalled
in~\cite{dvanhorn:Bloch2000Scheme}):
%
\begin{enumerate}
\item \emph{introduce only those language constructs that are necessary to
  teach programming principles}, and
\item \emph{choose a language with as
  few language constructs as possible, and one in which they can be
  introduced one at a time}.
\end{enumerate}

@@

At Northeastern, the College of Computer \& Information Science (CCIS)
requires a four course introductory sequence in the first year.  The
first semester features both a course on discrete mathematics and
 an introduction to programming following the \emph{How to Design
  Programs} curriculum.  The second semester follows with a course on
object-oriented programming and one featuring formal reasoning about
programs, both on paper and with the ACL2 theorem
prover~\cite{dvanhorn:Kaufmann2000ComputerAided}.

@@

This combination sets the constraints under which we designed our
approach.  Our students begin the course with a firm grasp of
data-driven program design, as well as experience with the tools used
in the first course:
